---
title: Run historical R versions on today's computers
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Experimental!

Note that these containers are currently experimental!

Found a bug?
[Please let us now!](https://github.com/r-hub/evercran/issues/new)
Thanks!

## TL;DR

```sh
docker pull ghcr.io/r-hub/evercran/1.0.0
docker run -ti ghcr.io/r-hub/evercran/1.0.0
```

```
R : Copyright 2000, The R Development Core Team
Version 1.0.0  (February 29, 2000)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type	"?license" or "?licence" for distribution details.

R is a collaborative project with many contributors.
Type	"?contributors" for a list.

Type	"demo()" for some demos, "help()" for on-line help, or
    	"help.start()" for a HTML browser interface to help.
Type	"q()" to quit R.

>
```

## R 0.0 (alpha-test) -- R 2.5.1

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 3.1 (Sarge). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work because of certificate errors.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran sarge main' \
    >> /etc/apt/sources.list
  apt-get update -y
  apt-get install -y r-1.0.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v1.0.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)
* If you run this image on `x86_64` machines, then `arch`, `uname -a` and
  potentially similar commands report the architecture incorrectly as
  `x86_64`. Use the `linux32` entrypoint to work around this. We don't do
  this by default currently, because it causes issues on arm64 hosts
  (see https://github.com/r-hub/evercran/issues/3). E.g.:
  ```
  ❯ docker run -ti ghcr.io/r-hub/evercran/0.49:latest arch
  x86_64

  ❯ docker run -ti --entrypoint linux32 ghcr.io/r-hub/evercran/0.49:latest arch
  i686
  ```

## R 2.6.0 -- R 2.8.1

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 4.0 (Etch). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work because of they use TLS v1.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran etch main' \
    >> /etc/apt/sources.list
  curl -L http://ppa.r-pkg.org/rhub.gpg.key | apt-key add -
  apt-get update -y
  apt-get install -y r-2.6.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v2.6.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)
* The entry point of the containers is `linux32` to make sure that all
  programs know that they are running on `i386`. Call `linux32` manually
  if you change the entry point.

## R 2.9.0 -- R 2.12.2

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 5.0.10 (Lenny). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work because of they use TLS v1.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran lenny main' \
    >> /etc/apt/sources.list
  curl -L http://ppa.r-pkg.org/rhub.gpg.key | apt-key add -
  apt-get update -y
  apt-get install -y r-2.9.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v2.9.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)
* The entry point of the containers is `linux32` to make sure that all
  programs know that they are running on `i386`. Call `linux32` manually
  if you change the entry point.

## R 2.13.0 -- R 2.15.3

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 6.0.10 (Squeeze). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work because of they use TLS v1.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran squeeze main' \
    >> /etc/apt/sources.list
  curl -L http://ppa.r-pkg.org/rhub.gpg.key | apt-key add -
  apt-get update -y
  apt-get install -y r-2.13.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v2.13.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)
* The entry point of the containers is `linux32` to make sure that all
  programs know that they are running on `i386`. Call `linux32` manually
  if you change the entry point.

## R 3.0.0 -- R 4.3.1

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 12.1 (Bookworm). Notes:

* All containers are available with `linux/amd64` and `linux/arm64` 
  architectures.
* The amd64 Debian packages are available from
  https://github.com/rstudio/r-builds.
* The arm64 DEbian packages are available from the GitHub Releases at the
  https://github.com/r-hub/R repository E.g. 
  https://github.com/r-hub/R/releases/tag/v4.3.1.
* To install additional R versions (or to install in a plain
  `debian:12` container, download it with `wget` or `curl`, ant install it
  with `apt install ./<filename>`. E.g.:
  ```sh
  apt-get update
  apt-get install -y curl
  curl -LO https://github.com/r-hub/R/releases/download/v4.3.1/r-rstudio-debian-12-4.3.1_1_arm64.deb
  apt install -y ./r-rstudio-debian-12-4.3.1_1_arm64.deb
  /opt/R/4.3.1/bin/R
  ```

  


## Containers with multiple R versions

For convenience we have containers that include many R versions installed.

All R versions are installed in `/opt/R/<version>`, so you can run one with
```
/opt/R/<version>/bin/R
```
For example to run R 0.65.1:
```
/opt/R/0.65.1/bin/R
```

| R versions                             | Container                    | OS                      |
|:---------------------------------------|:-----------------------------|:------------------------|
| R 0.0 (alpha-test) -- R 0.16.1 (alpha) | `ghcr.io/r-hub/evercran/pre` | Debian Sarge 3.1 i386   |
| R 0.49  -- R 1.0.0                     | `ghcr.io/r-hub/evercran/0.x` | Debian Sarge 3.1 i386   |
| R 1.0.0 -- R 1.9.1                     | `ghcr.io/r-hub/evercran/1.x` | Debian Sarge 3.1 i386   |
| R 2.0.0 -- R 2.15.3                    | `ghcr.io/r-hub/evercran/2.x` | Debian Wheezy 7.11 i386 |

## List of all containers

```{r, results = "asis", echo = FALSE}
writeLines("| R version | Container | Platform | OS |")
writeLines("|:----------|:----------|:---------|:---|")
rv <- readLines("containers/prerelease-versions.txt")
lab <- paste(rv, "(alpha)")
lab[1] <- "0.0 (alpha-test)"
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Sarge 3.1 i386 |",
  format(lab), format(rv)
))
rv <- readLines("containers/sarge-versions.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Sarge 3.1 i386 |",
  format(rv), format(rv)
))
rv <- readLines("containers/etch-versions.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Etch 4.0r9 i386 |",
  format(rv), format(rv)
))
rv <- readLines("containers/lenny-versions.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Lenny 5.0.10 i386 |",
  format(rv), format(rv)
))
rv <- readLines("containers/squeeze-versions.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Squeeze 6.0.10 i386 |",
  format(rv), format(rv)
))
rv <- readLines("containers/bookworm-versions.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/amd64`, `linux/arm64`| Debian bookworm 12.1 |",
  format(rv), format(rv)
))
```

## Similar projects, inspiration

* [rcheology](https://github.com/hughjonesd/rcheology) --
Data on Base Packages for Previous Versions of R
* [rang](https://github.com/chainsawriot/rang) --
(Re)constructing R computational environments

## Thanks!

* [debian/eol](https://hub.docker.com/r/debian/eol/) --
End of Life Debian versions (pointing at archive.debian.org)
* Posit's https://github.com/rstudio/r-builds project.

## License

See https://www.r-project.org/Licenses/ for the R licenses

The Dockerfiles and tools in this repository are licensed under the
MIT License.

© [R Consortium](https://github.com/rconsortium)
