---
title: Run historical R versions on today's computers
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Experimental!

Note that these containers are currently experimental!

Found a bug?
[Please let us know!](https://github.com/r-hub/evercran/issues/new)
Thanks!

## TL;DR

```sh
docker pull ghcr.io/r-hub/evercran/1.0.0
docker run -ti ghcr.io/r-hub/evercran/1.0.0
```

```
R : Copyright 2000, The R Development Core Team
Version 1.0.0  (February 29, 2000)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type	"?license" or "?licence" for distribution details.

R is a collaborative project with many contributors.
Type	"?contributors" for a list.

Type	"demo()" for some demos, "help()" for on-line help, or
    	"help.start()" for a HTML browser interface to help.
Type	"q()" to quit R.

>
```

## The evercran CRAN snapshots

The evercran project has daily CRAN snapshots, starting from the very
beginning up to today. The URLs are of the form
```
http://evercran.r-pkg.org/YYYY/MM/DD
https://evercran.r-pkg.org/YYYY/MM/DD
```
and they start at 1999/01/01.

Package downloads are redirected to CRAN, so you will need an R download
method that follows redirection.

All the evercran containers (from R 0.65.1) in this repository are already
set up to an appropriate evercran snapshot: they use the date of the
*next* R release. (Except for the container of the latest R release, of
course, that uses a regular CRAN mirror instead of a snapshot.)

To use an evercran snapshot on a different R installation,
this should work on all R versions, as of today:
```
options(CRAN = "http://evercran.r-pkg.org/2010/10/10")
options(repos = c(CRAN = "http://evercran.r-pkg.org/2010/10/10"))
options(download.file.method = "wget")
```
Modify the URLs to use the snapshot date you'd like.
Add this to the `~/.Rprofile` file.
You'll also need to install `wget` for the downloads to
work.
You can use a https URL if your `wget` installation supports https.

## R 0.0 (alpha-test) -- R 2.5.1

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 3.1 (Sarge). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work because of certificate errors.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran sarge main' \
    >> /etc/apt/sources.list
  apt-get update -y
  apt-get install -y r-1.0.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v1.0.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)

## R 2.6.0 -- R 2.8.1

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 4.0 (Etch). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work well, because they use TLS v1, which is not supported by many
  servers.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran etch main' \
    >> /etc/apt/sources.list
  curl -L http://ppa.r-pkg.org/rhub.gpg.key | apt-key add -
  apt-get update -y
  apt-get install -y r-2.6.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v2.6.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)

## R 2.9.0 -- R 2.12.2

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 5.0.10 (Lenny). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work well, because they use TLS v1, which is not supported by many
  servers.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran lenny main' \
    >> /etc/apt/sources.list
  curl -L http://ppa.r-pkg.org/rhub.gpg.key | apt-key add -
  apt-get update -y
  apt-get install -y r-2.9.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v2.9.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)

## R 2.13.0 -- R 2.15.3

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 6.0.10 (Squeeze). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work well, because they use TLS v1, which is not supported by many
  servers.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran squeeze main' \
    >> /etc/apt/sources.list
  curl -L http://ppa.r-pkg.org/rhub.gpg.key | apt-key add -
  apt-get update -y
  apt-get install -y r-2.13.0
  ```
* The Debian R packages are also available as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v2.13.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)

## R 3.0.0 -- R 4.4.2

```sh
docker pull ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 12.1 (Bookworm). Notes:

* All containers are available with `linux/amd64` and `linux/arm64`
  architectures.
* The amd64 Debian packages are available from
  https://github.com/rstudio/r-builds.
* The arm64 Debian packages are available from the GitHub Releases at the
  https://github.com/r-hub/R repository E.g.
  https://github.com/r-hub/R/releases/tag/v4.3.1.
* To install additional R versions (or to install in a plain
  `debian:12` container), download it with `wget` or `curl`, and install it
  with `apt install ./<filename>`. E.g.:
  ```sh
  apt-get update
  apt-get install -y curl
  curl -LO https://github.com/r-hub/R/releases/download/v4.3.1/r-rstudio-debian-12-4.3.1_1_arm64.deb
  apt install -y ./r-rstudio-debian-12-4.3.1_1_arm64.deb
  /opt/R/4.3.1/bin/R
  ```




## Containers with multiple R versions

For convenience we have containers that include many R versions installed.

All R versions are installed in `/opt/R/<version>`, so you can run one with
```
/opt/R/<version>/bin/R
```
For example to run R 0.65.1:
```
/opt/R/0.65.1/bin/R
```

| R versions                             | Container                    | OS                      |
|:---------------------------------------|:-----------------------------|:------------------------|
| R 0.0 (alpha-test) -- R 0.16.1 (alpha) | `ghcr.io/r-hub/evercran/pre` | Debian Sarge 3.1 i386   |
| R 0.49  -- R 1.0.0                     | `ghcr.io/r-hub/evercran/0.x` | Debian Sarge 3.1 i386   |
| R 1.0.0 -- R 1.9.1                     | `ghcr.io/r-hub/evercran/1.x` | Debian Sarge 3.1 i386   |
| R 2.0.0 -- R 2.15.3                    | `ghcr.io/r-hub/evercran/2.x` | Debian Wheezy 7.11 i386 |

## List of all containers

```{r, results = "asis", echo = FALSE}
writeLines("| R version | Container | Platform | OS |")
writeLines("|:----------|:----------|:---------|:---|")
rv <- readLines("containers/versions-prerelease.txt")
lab <- paste(rv, "(alpha)")
lab[1] <- "0.0 (alpha-test)"
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Sarge 3.1 |",
  format(lab), format(rv)
))
rv <- readLines("containers/versions-sarge.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Sarge 3.1 |",
  format(rv), format(rv)
))
rv <- readLines("containers/versions-etch.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Etch 4.0r9 |",
  format(rv), format(rv)
))
rv <- readLines("containers/versions-lenny.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Lenny 5.0.10 |",
  format(rv), format(rv)
))
rv <- readLines("containers/versions-squeeze.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Squeeze 6.0.10 |",
  format(rv), format(rv)
))
rv <- readLines("containers/versions-bookworm.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/amd64`, `linux/arm64`| Debian bookworm 12.1 |",
  format(rv), format(rv)
))
```

## Frequently asked questions

<details>
<summary>Why does `list.files()` return an empty vector?</summary>

If you see something like
```
> list.files("/opt/R")
character(0)
```
then you need to run the entry point of the container, see the question
after the next one.
</details>

<details>
<summary>
Why is the architecture detected as `x86_64` instead of `i386`?
</summary>

You need to run the entry point of the container, see the next question.
</details>

<details>
<summary>`docker run` works, but `docker exec` does not, why?</summary>

Running 32bit (`i386`) containers on arm64 and amd64 platforms is tricky,
and to make everything work properly, we need to run a special entry point.
This entry point runs with `docker run`, but it does not for `docker exec`,
and you need to run it manually:
```
docker exec -it <container> entrypoint.sh bash
```

If you don't run the entry point then on arm64 hosts you might get file
system errors (https://github.com/r-hub/evercran/issues/7). On amd64
platforms typically the architecture will be mis-detected
(https://github.com/r-hub/evercran/issues/3).
</details>

## Similar projects, inspiration

* [rcheology](https://github.com/hughjonesd/rcheology) --
Data on Base Packages for Previous Versions of R
* [rang](https://github.com/chainsawriot/rang) --
(Re)constructing R computational environments

## Thanks!

* [debian/eol](https://hub.docker.com/r/debian/eol/) --
End of Life Debian versions (pointing at archive.debian.org)
* Posit's https://github.com/rstudio/r-builds project.

## License

See https://www.r-project.org/Licenses/ for the R licenses

The Dockerfiles and tools in this repository are licensed under the
MIT License.

© [R Consortium](https://github.com/rconsortium)
