---
title: Run historical R versions on today's computers
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## TL;DR

```sh
docker pull --platform linux/i386 ghcr.io/r-hub/evercran/1.0.0
docker run -ti ghcr.io/r-hub/evercran/1.0.0
```

```
R : Copyright 2000, The R Development Core Team
Version 1.0.0  (February 29, 2000)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type	"?license" or "?licence" for distribution details.

R is a collaborative project with many contributors.
Type	"?contributors" for a list.

Type	"demo()" for some demos, "help()" for on-line help, or
    	"help.start()" for a HTML browser interface to help.
Type	"q()" to quit R.

>
```

## Containers with multiple R versions

All R versions are installed in `/opt/R/<version>`, so you can run one with
```
/opt/R/<version>/bin/R
```
For example to run R 0.65.1:
```
/opt/R/0.65.1/bin/R
```

| R versions        | Container                    | OS                    |
|:------------------|:-----------------------------|:----------------------|
| R 0.49 -- R 1.0.0 | `ghcr.io/r-hub/evercran/0.x` | Debian Sarge 3.1 i386 |

## All containers

```{r, results = "asis", echo = FALSE}
writeLines("| R version | Container | OS |")
writeLines("|:----------|:----------|:---|")
rv <- readLines("containers/sarge-versions.txt")
writeLines(sprintf(
  "| R %s | ghcr.io/r-hub/evercran/%s | Debian Sarge 3.1 i386 |", 
  format(rv), format(rv)
))
```

## Similar projects, inspiration

* [rcheology](https://github.com/hughjonesd/rcheology) --
Data on Base Packages for Previous Versions of R
* [rang](https://github.com/chainsawriot/rang) --
(Re)constructing R computational environments

## Thanks!

* [debian/eol](https://hub.docker.com/r/debian/eol/) --
End of Life Debian versions (pointing at archive.debian.org) 

## License

See https://www.r-project.org/Licenses/ for the R licenses

The Dockerfiles and tools in this repository are licensed under the
MIT License.

© [R Consortium](https://github.com/rconsortium)
