---
title: Run historical R versions on today's computers
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options:
  markdown:
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

## Experimental!

Note that these containers are currently experimental!

Found a bug?
[Please let us now!](https://github.com/r-hub/evercran/issues/new)
Thanks!

## TL;DR

```sh
docker pull --platform linux/i386 ghcr.io/r-hub/evercran/1.0.0
docker run -ti ghcr.io/r-hub/evercran/1.0.0
```

```
R : Copyright 2000, The R Development Core Team
Version 1.0.0  (February 29, 2000)

R is free software and comes with ABSOLUTELY NO WARRANTY.
You are welcome to redistribute it under certain conditions.
Type	"?license" or "?licence" for distribution details.

R is a collaborative project with many contributors.
Type	"?contributors" for a list.

Type	"demo()" for some demos, "help()" for on-line help, or
    	"help.start()" for a HTML browser interface to help.
Type	"q()" to quit R.

>
```

## R 0.49 -- R 1.9.1

```sh
docker pull --platform linux/i386 ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 3.1 (Sarge). Notes:

* All containers use the `linux/i386` architecture.
* While `wget` and `curl` have HTTPS support, in practice HTTPS
  does not work because of certificate errors.
* You can use the HTTP (not HTTPS!) PPA at https://ppa.r-pkg.org/evercran/
  to install more R versions:
  ```sh
  echo 'deb http://ppa.r-pkg.org/evercran sarge main' \
    >> /etc/apt/sources.list
  apt-get update -y
  apt-get install -y r-1.0.0
  ```
* The Debian R packages are also availablr as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v1.0.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)
* The entry point of the containers is `linux32` to make sure that all
  programs know that they are running on `i386`. Call `linux32` manually
  if you change the entry point.

## R 2.0.0 -- R 2.15.3

```sh
docker pull --platform linux/i386 ghcr.io/r-hub/evercran/<version>
docker run -ti ghcr.io/r-hub/evercran/<version>
```

These containers use Debian 7.11 (Wheezy). Notes:

* All containers use the `linux/i386` architecture.
* `wget` and `curl` have HTTPS support, but you'll need to install the
  `ca-certificates` package (it is automatically installed if you install
  `curl`). This triggers a certificate update. Not all HTTPS web sites
  will work, probably, because of how old versions of the SSL libraries
  handle alternate vertificate chains, or other reasons.
* The Debian R packages are availablr as GitHub Releases, in the
  https://github.com/r-hub/R repository, e.g.:
  https://github.com/r-hub/R/releases/tag/v2.0.0
  (But you cannot directly download these from the containers, because
  of the broken HTTPS. Use the PPA to download them.)
* To install additional R versions (or to install in a plain
  `debian/eol:wheezy` container, download it with `wget` or `curl`, install it
  with `dpkg -i` and then call `apt-get -f install` to also install its
  dependencies. E.g.:
  ```sh
  apt-get update
  apt-get install -y wget ca-certificates
  wget https://github.com/r-hub/R/releases/download/v2.0.0/r-evercran-debian-7.11-2.0.0_1-1_i386.deb
  dpkg -i r-evercran-debian-7.11-2.0.0_1-1_i386.deb || true
  apt-get -fy install
  /opt/R/2.0.0/bin/R
  ```
* The entry point of the containers is `linux32` to make sure that all
  programs know that they are running on `i386`. Call `linux32` manually
  if you change the entry point.

## Containers with multiple R versions

For convenience we have containers that include many R versions installed.

All R versions are installed in `/opt/R/<version>`, so you can run one with
```
/opt/R/<version>/bin/R
```
For example to run R 0.65.1:
```
/opt/R/0.65.1/bin/R
```

| R versions          | Container                    | OS                      |
|:--------------------|:-----------------------------|:------------------------|
| R 0.49  -- R 1.0.0  | `ghcr.io/r-hub/evercran/0.x` | Debian Sarge 3.1 i386   |
| R 1.0.0 -- R 1.9.1  | `ghcr.io/r-hub/evercran/1.x` | Debian Sarge 3.1 i386   |
| R 2.0.0 -- R 2.15.3 | `ghcr.io/r-hub/evercran/2.x` | Debian Wheezy 7.11 i386 |

## List of all containers

```{r, results = "asis", echo = FALSE}
writeLines("| R version | Container | Platform | OS |")
writeLines("|:----------|:----------|:---------|:---|")
rv <- readLines("containers/sarge-versions.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian Sarge 3.1 i386 |",
  format(rv), format(rv)
))
rv <- readLines("containers/wheezy-versions.txt")
writeLines(sprintf(
  "| R %s | `ghcr.io/r-hub/evercran/%s` | `linux/i386` | Debian wheezy 7.11 i386 |",
  format(rv), format(rv)
))
```

## Similar projects, inspiration

* [rcheology](https://github.com/hughjonesd/rcheology) --
Data on Base Packages for Previous Versions of R
* [rang](https://github.com/chainsawriot/rang) --
(Re)constructing R computational environments

## Thanks!

* [debian/eol](https://hub.docker.com/r/debian/eol/) --
End of Life Debian versions (pointing at archive.debian.org)

## License

See https://www.r-project.org/Licenses/ for the R licenses

The Dockerfiles and tools in this repository are licensed under the
MIT License.

© [R Consortium](https://github.com/rconsortium)
